---
title: "EDA"
author: "Xiaoyu Wu"
date: "2023-10-04"
output: github_document
---
```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width=6,
  fig.asp= .6,
  out.width="90%"
)

theme_set(theme_minimal()+theme(legend.position="bottom"))

options(
  ggplot2.cotinuous.color="viridis",
  ggplot2.cotinuous.fill="viridis"
)
# overwrite ggplot default coloring and take care of call the colors and fills 

scale_color_discrete=scale_color_viridis_d
scale_fill_discrete=scale_fill_viridis_d
```

## Load Dataset 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) |>
  select(name, id, everything())
weather_df
```
## Initial Numeric Exploitation
Histogram of precipitation 
```{r}
weather_df |> 
  ggplot(aes(x = prcp)) + 
  geom_histogram()
```
Examine the relatively few days have very high precipitation
```{r}
weather_df |> 
  filter(prcp >= 1000)
```
Scatterplot 
```{r}
weather_df |> 
  filter(tmax >= 20, tmax <= 30) |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + 
  geom_point(alpha = .75)
```
## `Group by`
```{r}
weather_df |>
  group_by(name, month) |>
  
  ungroup()
```
## Counting Things: count the number of observations in each month
```{r}
weather_df |>
  group_by(month) |>
  summarize(n_obs = n())
# n() is the function and n_obs is the new column added 
```
Summarize: count the number of observations in each month and name
```{r}
weather_df |>
  group_by(name, month) |>
  summarize(n_obs = n())
```
count() in place of group_by()
```{r}
weather_df |>
  count(month, name = "n_obs")
# n_obs is the new column added 
```
other helpful counters 
```{r}
weather_df |>
  group_by(month) |>
  summarize(
    n_obs = n(),
# count the number of observations in each month 
    n_days = n_distinct(date))
# count the number of distinct values of date in each month
```

## 2 times 2 tables 
```{r}
weather_df |> 
  filter(name != "Waikiki_HA") |>
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) |> 
  group_by(name, cold) |> 
  summarize(count = n())
```
re-organize into a more standard (non-tidy) 2x2 table 
```{r}
weather_df |> 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) |> 
  filter(name != "Waikiki_HA") |> 
  janitor::tabyl(name, cold)
```
## General summaries
```{r}
weather_df |>
  group_by(month) |>
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
# get rid of the missing values by using "na.rm = TRUE"
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))
```
You can group by more than one variable.
```{r}
weather_df |>
  group_by(name, month) |>
  summarize(
    mean_tmax = mean(tmax),
    median_tmax = median(tmax))
```
suppose you want to summarize many columns  
```{r}
weather_df |>
  group_by(name, month) |>
  summarize(across(tmin:prcp, mean))
```
we can take create a plot based on the monthly summary
```{r}
weather_df |>
  group_by(name, month) |>
  summarize(mean_tmax = mean(tmax)) |>
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```
pivot wider format
```{r}
weather_df |>
  group_by(name, month) |>
  summarize(mean_tmax = mean(tmax)) |> 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) |> 
  knitr::kable(digits = 1)
# retain how many digits after zero
```
## `grouped by` and `mutate`
compare the daily max temperature to the annual average max temperature for each station separately
```{r}
weather_df |>
  group_by(name) |>
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) |> 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```
## Window functions
we can find the max temperature ranking within month.
```{r}
weather_df |>
  group_by(name, month) |>
  mutate(temp_ranking = min_rank(tmax))
# rank from lowest to highest 
```
keep only the day with the lowest max temperature within each month
```{r}
weather_df |>
  group_by(name, month) |>
  filter(min_rank(tmax) == 1)
```
keep the three days with the highest max temperature
```{r}
weather_df |>
  group_by(name, month) |>
  filter(min_rank(desc(tmax)) < 4)
```
find the day-by-day change in max temperature within each station over the year
```{r}
weather_df |>
  group_by(name) |>
  mutate(temp_change = tmax - lag(tmax))
```

```{r}
weather_df |>
  group_by(name) |>
  mutate(temp_change = tmax - lag(tmax)) |>
  summarize(
    temp_change_sd = sd(temp_change, na.rm = TRUE),
    temp_change_max = max(temp_change, na.rm = TRUE))
```
## learning assessment 
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") |>
  select(id, visit, everything()) |>
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) |>
  arrange(id, visit)
pulse_data
pulse_data |> 
  group_by(visit) |> 
  summarize(
    mean_bdi = mean(bdi, na.rm = TRUE),
    median_bdi = median(bdi, na.rm = TRUE)) |> 
  knitr::kable(digits = 3)
```

## Quick note
summarize only gets you so far 
