---
title: "Visualization 2"
author: "Xiaoyu Wu"
date: "2023-09-29"
output: html_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```
Load Dataset 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```
## Remember this plot...? 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color=name)) + 
  geom_point( alpha = .5)
```
## Add labels 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color=name)) + 
  geom_point( alpha = .5) + 
  labs(
    title = "Temperature plot",
# specify the title 
    x = "Minimum daily temperature (C)",
# specify the x axis label  
    y = "Maxiumum daily temperature (C)",
# specify the y axis label 
    caption = "Data from the rnoaa package; temperatures in 2017"
  )
# add a caption 
```
## Scales 
Start with the same plot; x and y scales.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบ C", "0", "15")) +
# scale label 
# categorical or factor variable we use scale_x_discrete 
    scale_y_continuous(
    trans = "sqrt", 
# scale_y_sqrt() can be added to a ggplot object to transform the Y scale
  # OR trans="log"
    position = "right")
# change the scale position to the right 
```

Look at color scales 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,color=name)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017") + 
    scale_color_hue(
      name = "Location", 
# change the title of the caption from "name" to "location" 
      h = c(100, 300))
# arguments to scale_color_hue() control the color scale 
```
OR 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,color=name)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017") + 
    viridis::scale_color_viridis(
      name = "Location",
      discrete=TRUE)
# viridis::scale_color_viridis assume the scale is continuous, we have to tell it's discrete
```

Themes

Shift the legend 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,color=name)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017") + 
    viridis::scale_color_viridis(
      name = "Location",
      discrete=TRUE) + 
  theme(legend.position = "bottom")
# change the position of legend to bottom 
```

theme built in 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,color=name)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package; temperatures in 2017") + 
    viridis::scale_color_viridis(
      name = "Location",
      discrete=TRUE) + 
      theme_bw()+
      theme(legend.position="bottom")
# theme black and white
# theme_minimal() 
# theme_classic()
# ggthemes::theme_economist()
# ggthemes::theme_excel()
```

learning assessment
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) + 
  geom_point(aes(size = prcp), alpha = .75) + 
  labs(
    title = "Temperature plot",
    x = "Date",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE) + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```

## Setting options on the very beginning of rmarkdown document 

```{r, eval=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width=6,
  fig.asp= .6,
  out.width="90%"
)

theme_set(theme_minimal()+theme(legend.position="bottom"))

options(
  ggplot2.cotinuous.color="viridis",
  ggplot2.cotinuous.fill="viridis"
)
# overwrite ggplot default coloring and take care of call the colors and fills 

scale_color_discrete=scale_color_viridis_d
scale_fill_discrete=scale_fill_viridis_d
```


## data argument in `geom`

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")
# split weather_df into separate datasets for Central Park and Waikiki
ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
# use one in the ggplot() call and another in geom_line()
```

## patchwork

remember faceting?
```{r}
weather_df |> 
  ggplot(aes(x = tmin,fill=name)) +
# fill only applies to density plot 
  geom_density(alpha = .5) +
# adjust is about the same with binwidth 
  facet_grid(.~name)
```

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```
## data manipulation 

Control your factors. 
```{r}
weather_df  %>%
  mutate(
    name=factor(name),
  # character to fatcor version of name
    name=forcats::fct_relevel(name,c("Waikiki_HA"))
  ) %>%
# change the order of name 
  ggplot(aes(x=name,y=tmax,fill=name)) +
  geom_violin(alpha=.5)
```

What if I wanted densities for tmin and tmax simultaneouly? 

```{r}
weather_df  %>%
  filter(name=="CentralPark_NY") %>%
  pivot_longer(
    tmax:tmin,
    names_to="observations",
    values_to="temperatures"
  ) %>%
  ggplot(aes(x=temperatures,fill=observations))+
  geom_density(alpha=.5)
```

## Revisit the pups 

Data from the FAS study. 
```{r}
pup_data=
  read_csv("./data/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  mutate(sex=recode(sex,`1`="male",`2`="female"))

litters_data=
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  separate(group,into=c("dose","day_of_tx"),sep=3)

fas_data=left_join(pup_data,litters_data,by="litter_number")

fas_data %>%
  select(dose,day_of_tx,starts_with("pd_")) %>%
  pivot_longer(
    pd_ears:pd_walk,
    names_to="outcome",
    values_to="pn_day"
  )  %>%
  drop_na() %>%
  mutate(outcome=forcats::fct_relevel(outcome,"pn_ears","pn_pivot","pn_walk","pn_eyes")) %>%
# factor manipulation 
  ggplot(aes(x=dose,y=pn_day))+
  geom_violin()+
  facet_grid(day_of_tx~outcome)

```

```{r}
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
# reorders name according to tmax values in each name
```

