---
title: "Visualization 1"
author: "Xiaoyu Wu"
date: "2023-09-27"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## R Markdown

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
weather_df
```

## Let's make a plot! 
Create my first scatterplot ever 
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```
New approach, same plot, we do this way more often!  
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```
Save and edit a plot object 
```{r}
ggp_weather = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) 

ggp_weather + geom_point()
# Save the output of ggplot() to an object and modify / print it later.
```

## Advanced scatterplot

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax,color=name)) + 
  geom_point()
# Incorporate using the color aesthetic.
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax,color=name)) + 
# color applies to all lines below 
  geom_point() +
  geom_smooth(se = FALSE)
# Add a smooth curve
```
what about the `aes` placement 
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color=name)) +
# color only applies to the scatterplot
  geom_smooth(se = FALSE)
# mostly we cannot define asthetics in the first step, we prefer to define in geom_point so that the color is not everywhere 
```
Let's facet some things! 
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax,color=name)) + 
  geom_point(alpha = .3) +
# alpha makes individual points more transparent, 30% transparent and 70% solid.  
  geom_smooth(se = FALSE,size=2)+
# change the size of the smooth line 
  facet_grid(. ~ name)
# Add facet based on name and seperate on "y axis"
```
Let's combine some elements and try a new plot. 
```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax,color=name)) + 
  geom_point(aes(size=prcp),alpha = .3) +
  geom_smooth(se = FALSE)+
  facet_grid(. ~ name)
# Set x axis as date and show the precipitation on graph 
```
Try assigning specific color 
```{r}
weather_df |>
  filter(name=="CentralPark_NY") |>
  ggplot(aes(x = date, y = tmax)) + 
  geom_point(color="blue")
```

Learning Assessment 
```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32) |> 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)
# Focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth(). 
```
## Some small notes

How many geoms have to exist?
You can have whatever geoms you want. 
You can use a neat geom!
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax,color=name)) + 
  geom_hex() 
# geom_bin2d()
# geom_density2d()
```

## Odds and Ends 

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 
```

```{r}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")
# Defining the color of the points by hand
```

```{r}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))
# Creating a color variable that has the value blue everywhere; ggplot is then assigning colors according to this variable using the default color scheme.
```

## Univariate plots
Histograms are really great.
```{r}
weather_df |> 
   ggplot(aes(x = tmin)) + 
  geom_histogram()
# histogram
```
Can we add color..

```{r}
weather_df |> 
  ggplot(aes(x = tmin,fill=name)) + 
  geom_histogram(position="dodge") +
# the bins sit next to each other 
  facet_grid(.~name)
```
Let's try a new geometry!
```{r}
weather_df |> 
  ggplot(aes(x = tmin,fill=name)) + 
  geom_density(alpha = .4,adjust = .5) +
  facet_grid(.~name)
# Density plot
```

What about boxplots?

```{r}
weather_df |> 
  ggplot(aes(x = name,y=tmin)) + 
  geom_boxplot()
# boxplot
```
Trendy plots :)
```{r}
weather_df |> 
  ggplot(aes(x = name,y=tmin,fill=name)) + 
  geom_violin(alpha=0.5) + 
  stat_summary(fun="median")
# violin plots
```
Ridge plots
```{r}
weather_df |> 
  ggplot(aes(x = tmin,y=name)) + 
  geom_density_ridges() 
# ridge plots
```
## Save and Export 
```{r}
weather_plot =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax,color=name)) + 
  geom_point(alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)
```
What about emdedding? 
```{r,fig.width=6,fig.height=3}
weather_plot
# embed figures in r-markdown document 
```
Embed at different size. 
```{r, fig.width=12,fig.height=6}
weather_plot
# embed figures in r-markdown document 
```

